Longitudinal Principal Manifold Estimation

### 2023-07-12 paper/main.tex
update simulation description

A factorial design is used to run the simulation studies, with the factor levels set as follows:
\begin{itemize}
  \item $\alpha, \beta, \zeta \in \left\{0, 0.5, 1.0, 1.5, 2.0\right\}$
  \item Study duration: $\left\{1, 2, 5, 10\right\}$
  \item Interval between images: $\left\{0.1, 0.25, 0.5, 1\right\}$
  \item Sample size: $\left\{1000, 10000\right\}$
\end{itemize}

Thus, each embedding map is simulated a total of 16,000 times. The results of these simulations are summarized in Table \ref{table:simulation_results}. All code required to reproduce the simulation results and the associated figures is provided at \texttt{https://github.com/rjzielinski/lpme-project}.

Visualizations of the results from some of the simulation cases, truncated for concision to include only integer-valued time points, are shown in Figures \ref{fig:sim_case1}, \ref{fig:sim_case4} and \ref{fig:sim_case7}. Informally, the performance of each estimation method can be observed by considering the proximity of each method's estimated manifold to the true manifold (shown in red) over time. Thus, in Figure \ref{fig:sim_case1}, we see that the LPME-estimated manifold bears the closest resemblance to the true manifold underlying the data, while the manifolds estimated by PME and the principal curve method show greater responses to temporary random fluctuations in the data at each time point.

@@

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & & & & & & & & & \\
    LPME & & & & & & & & & \\
    PME & & & & & & & & & \\
    PC/PS & & & & & & & & & \\
    \hline
  \end{tabular}
  \caption{MSD Comparison.}
  \label{table:simulation_results}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sim_case8}
  \caption{Simulation Case 7}
  \label{fig:sim_case7}
\end{figure}

\zielinski{A brief discussion of the simulation results will follow when they become available.}

>>>>

A factorial design is used to run the simulation studies, with the factor levels set as follows:
\begin{itemize}
  \item $\alpha, \beta, \zeta \in \left\{0, 0.1, 0.25, 0.5, 1.0, 2.0\right\}$
  \item Study duration: $\left\{1, 2, 5, 10\right\}$
  \item Interval between images: $\left\{0.1, 0.25, 0.5, 1\right\}$
  \item Sample size: $\left\{1000, 10000\right\}$
  \item Longitudinal change model: Constant, Linear, Quadratic, Sinusoidal
\end{itemize}

Thus, each embedding map is simulated a total of 27,648 times. All code required to reproduce the simulation results and the associated figures is provided at \texttt{https://github.com/rjzielinski/lpme-project}.

Visualizations of the results from some of the simulation cases, truncated for concision to include only integer-valued time points, are shown in Figures \ref{fig:sim_case1}, \ref{fig:sim_case4} and \ref{fig:sim_case6}. Informally, the performance of each estimation method can be observed by considering the proximity of each method's estimated manifold to the true manifold (shown in red) over time. Thus, in Figure \ref{fig:sim_case1}, we see that the LPME-estimated manifold bears the closest resemblance to the true manifold underlying the data, while the manifolds estimated by PME and the principal curve method show greater responses to temporary random fluctuations in the data at each time point.

@@

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.758 (1.28) & 0.884 (0.726) & N/A & 198 (466) & 1.91 (1.76) & 13.4 (51.4) & N/A & N/A & N/A \\
    LPME & 4.55 (33.4) & 0.483 (0.471) & N/A & 198 (843) & 1.45 (1.34) & 8.99 (31.7) & N/A & N/A & N/A \\
    PME & 1.01 (2.01) & 3.38 (12.1) & N/A & 194 (461) & 4.49 (6.26) & 12.4 (41.9) & N/A & N/A & N/A \\
    PC/PS & 0.688 (1.23) & 0.544 (0.276) & N/A & 143 (379) & 1.37 (1.26) & 5.58 (19.6) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison, Mean (SD).}
  \label{table:simulation_results_mean}
\end{table}

\zielinski{With results still incomplete, it seems like the worst LPME performance is concentrated in cases where the structure of the manifold changes drastically between time points. I have included these cases in the simulations to test where the method breaks down. For reporting purposes, would it be worth it to limit ourselves to more realistic cases without quite so much noise? Also, would it be worthwhile to limit the number of simulation cases we are considering for a more focused assessment?}

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.244 (0.576) & 0.784 (0.585) & N/A & 14.3 (84.0) & 1.63 (1.85) & 0.870 (4.49) & N/A & N/A & N/A \\
    LPME & 0.135 (0.265) & 0.406 (0.337) & N/A & 8.30 (35.6) & 1.21 (1.18) & 0.802 (2.85) & N/A & N/A & N/A \\
    PME & 0.351 (0.814) & 0.980 (1.61) & N/A & 12.3 (65.9) & 2.50 (3.56) & 0.883 (4.17) & N/A & N/A & N/A \\
    PC/PS & 0.201 (0.554) & 0.529 (0.215) & N/A & 7.66 (41.2) & 1.15 (1.13) & 0.867 (1.90) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison, Median (IQR).}
  \label{table:simulation_results_median}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sim_case7}
  \caption{Simulation Case 6}
  \label{fig:sim_case6}
\end{figure}

The mean, median, and standard deviation of the mean squared distance from the true underlying manifold values for the estimates of each approach, as well as for the data itself, are shown in Tables \ref{table:simulation_results_mean} and \ref{table:simulation_results_median}.
